
# Tidy {#tidy}

Reshaping the data is an important (if necessary) step to exploratory data analysis and preparatory data cleaning for modeling or creating specialized visualization. Further concepts about *tidy data* can be found in this paper [Tidy Data](http://vita.had.co.nz/papers/tidy-data.pdf).

The principles of tidy data are:

1. Each variable forms a column
2. Each observation forms a row
3. Each type of observational unit forms a table

* [`tidyr`](https://cran.r-project.org/web/packages/tidyr/index.html): An evolution of 'reshape2'. It's designed specifically for data tidying (not general reshaping or aggregating) and works well with 'dplyr' data pipelines. Tidy data complements R’s vectorized operations. R will automatically preserve observations as you manipulate variables. No other format works as intuitively with R.

**Examples: **

```{r, message=FALSE, warning=FALSE}
library(tidyr)
library(insuranceData)
library(magrittr)
data("AutoBi")
head(AutoBi)

# create an interaction column
g <- AutoBi %>%  unite(col = MARITAL_CLMAGE, MARITAL, CLMAGE, sep = "*")

head(g)

```


**Examples: ** 

Not all data that needs to be *tidied* comes in "long" format (i.e. the `spread()` function), so this data set below is put into tidy form using base R functions. This data set has its observations stored in the row names field, delimited with a semicolon ";". The original column contains a  mis-transformed value for the population density metric. The only variables' names are also a concatenation of the entire data set's column names delimited with a period ".".

* [Data dictionary](http://www.businessandeconomics.mq.edu.au/our_departments/Applied_Finance_and_Actuarial_Studies/acst_docs/glms_for_insurance_data/data/third_party_claims.txt)
* [Original source data used in the insuranceData package](http://www.businessandeconomics.mq.edu.au/our_departments/Applied_Finance_and_Actuarial_Studies/acst_docs/glms_for_insurance_data/data/Third_party_claims.xls)
```{r}
library(insuranceData)
library(magrittr)
data("Thirdparty")

head(Thirdparty) # rownames have column values
cols = strsplit(colnames(Thirdparty) , "." , fixed=T)   # a new vector to use later 
dput(unlist(cols))
rows2df <- sapply(rownames(Thirdparty), strsplit, ";", USE.NAMES = FALSE)

tidy_3PD <- data.frame(matrix(unlist(rows2df), nrow = nrow(Thirdparty), byrow=T))
colnames(tidy_3PD) <- c("lga", "sd", "claims", "accidents", "ki", "population",
"pop_density")
tidy_3PD$pop_density <- Thirdparty$lga.sd.claims.accidents.ki.population.pop_density / 1000000

head(tidy_3PD)

```

**Every variable forms a column and every observation forms a row which makes for a table!**

___________________
