# Visualization {#viz}

Data Visualization allows for the effective translation of complex models and processes to business applicable decisions. 

## **Exploratory Visualization**

Exploratory visualization invloves learning descriptive details prior to modeling efforts. Preemptive results from visualizing distributions can lead to more informed approachs in for example parameter tunning.

* [`ggplot2`](https://cran.r-project.org/web/packages/ggplot2/index.html): An implementation of the grammar of graphics in R. It combines the advantages of both base and lattice graphics: conditioning and shared axes are handled automatically, and you can still build up a plot step by step from multiple data sources. It also implements a sophisticated multidimensional conditioning system and a consistent interface to map data to aesthetic attributes. 

**Examples:**

```{r, ggplot2-exp1}
library(ggplot2)
library(insuranceData)
data("AutoClaims")
head(AutoClaims)
g <- ggplot(AutoClaims, aes(x = AGE)) +
            geom_bar() +
            facet_grid(. ~ GENDER)
g

```
_____________

```{r, ggplot2-exp2}
library(ggplot2)
library(insuranceData)


g2 <- ggplot(AutoClaims, aes(x = AGE, y = PAID, color = GENDER)) +
            geom_point() +
            geom_text(aes(label = STATE)) +
            theme_classic()
g2

```


## **Interactive Visualization**

* [`plotly`](https://cran.r-project.org/web/packages/plotly/index.html): Easily translate 'ggplot2' graphs to an interactive web-based version and/or create custom web-based visualizations directly from R. Once uploaded to a 'plotly' account, 'plotly' graphs (and the data behind them) can be viewed and modified in a web browser.

**Examples:**

```{r, message=FALSE, warning=FALSE}
suppressPackageStartupMessages(library(plotly))
library(insuranceData)
data("AutoCollision")
head(AutoCollision)

plot_ly(AutoCollision, x = Severity, y = Claim_Count, mode = "markers", 
        color = Severity, size = Severity)


```

## **Other visualization packages**

* [rcharts](http://ramnathv.github.io/rCharts/)
* [leaflet](https://rstudio.github.io/leaflet/)
* [ggvis](http://ggvis.rstudio.com/)
* [htmlwidget](http://www.htmlwidgets.org/)
* [googleVis](https://cran.r-project.org/web/packages/googleVis/vignettes/googleVis_examples.html)
